Discrepancy, Coresets, and Sketches in Machine Learning

### 2019-01-29 komlos.tex
some small edits

\section{Introduction}
In both machine learning and in streaming and sketching problems our goal is (often) to approximate sums or expectations of well behaved functions.
Specifically, we need to approximate $\E_x f(x)$ or $\frac{1}{n}\sum_{i=1}^{n} f(x_i)$ for every $f\in \F$ where $\F$ is a family of functions and $x_i \in \mathcal X$ are either sampled training examples or an arbitrary set of stream items. 
Standard generalization results show that for a large enough value of $n$ the average approximates the mean if the complexity of $\F$ is bounded and the samples $x_i$ are drawn i.i.d.\ from underlying distribution. We therefore focus on approximating the average, or rather, the sum $\sum_{i=1}^{n} f(x_i)$. For notational convenience we use a parameter $q \in \mathcal Q$ to index into $\F$ explicitly $F(q) = \sum_{i=1}^{n} f(x_i, q)$ \zk{It seems confusing that we are moving from $f\in \F$ to a universal single function $f$, $f(q,x)$. I'm thinking of calling it $g \in \F$ above and saying the for $\Q \equiv \F$, meaning there is a bijective mapping between them, we have $g(x)=f(q,x)$ for all $x$, for any matching pair $f,q$}. 
One should think about $q$ as either the model parameters or a query for the sketch.

The goal is to produce a coreset. This is a set of weights $w$ with at small number of non-zeros such that $\sum_{i=1}^{n}w_i f(x_i,q) := \tilde F(q)$ approximates $F(q)$.
Approximation here means that $|\tilde F(q)  - F(q)| \le \eps n$ either for all $q \in \mathcal Q$ simultaneously or for every fixed $q$ with probability at least $1-\delta$. 
There are more complicated formulations such as weak coresets which we will not touch upon in this manuscript. 
Generating a concise representation $\tilde F$ for $F$ allows one to optimize over $\tilde F$ instead of $F$ which is more efficient. 
Moreover, if the resulting coresets are mergeable, this could be done on separate streams without the need for communication or assuming randomness in the partitioning.

@@

\subsection{Coreset Complexity}
Here, we point out a direct connection between Coreset Complexity and Class Discrepancy.
The connection is a direct application of the folklore argument know as the ``the halving trick".

\begin{definition} [\bf Coreset Complexity] For a function $f:\X,\Q\rightarrow \R$ let $F(q) = \sum_{i=1}^{m} f(x_i,q)$ for a any set $\{x_1,\ldots,x_m\} \subset \X$.
For a set $S \subset [m]$ let $\tilde F(q) = \sum_{i \in S}w_i f(x_i,q)$ for some $w\in\R_+^m$ which is independent of $q$.
The Coreset Complexity of $f$ is the size of the smallest set $S$ such that $\forall q \in \Q \; |F(q)  - \tilde F(q)| \le \eps m$.
\end{definition}

\noindent The following facts are true for the common cases where $K_m = O(c/m)$ or $K_m = O(c/\sqrt{m})$.
\begin{fact}
Any function $f$ with Class Discrepancy $K_m(f) = O(c/m)$ has Coreset Complexity of $f$ $O(c/\eps)$.
The constant $c$ does not depend on $m$.
\end{fact}
\begin{fact}
Any function $f$ with Class Discrepancy $K_m(f) = O(c/\sqrt{m})$ has Coreset Complexity $O(c^2/\eps^2)$.
The constant $c$ does not depend on $m$.
\end{fact}
\begin{proof}
\noindent  Consider the signed sum of error function $E(q) = \sum_{i=1}^{m} \sigma_i f(x_i,q)$ where $\sigma_i \in \{-1,1\}$.
Now consider $\tilde F_{+}(q) = F(q) + E(q)   = \sum_{i | \sigma_i=1} 2 f(x_i,q)$ and similarly $\tilde F_{-}(q) = F(q) - E(q)  =  \sum_{i | \sigma_i=-1} 2 f(x_i,q)$. We have that both $\tilde F_{+}(q)$ and $\tilde F_{-}(q)$ are approximations for $F(q)$ obtained by coresets of item weights of $2$ and an error at most $|\tilde F_{\pm}(q)- F(q)| =  |E(q)|$. The above is true for any choice of signs $\sigma$, specifically, for those minimizing $\max_q | E(q)|$.
By Definition we can select signs such that $|E(q)| \le m K_m(f)$.

@@

\subsection{Streaming Coreset Complexity}\label{sec:sketch}
We claim that low Class Discrepancy implies concise streaming mergeable coresets as well. 
%
\begin{definition} [\bf Streaming Coreset Algorithm] 
A streaming coreset algorithm for $f:\X,\Q \rightarrow \R$ receives and arbitrary set $\{x_1,\ldots,x_m\} \subset \X$ one item after the other.
At time $t \le m$, the algorithm maintains a subset $S_t \subset \{x_1,\ldots,x_t\}$ and uses at most $O(|S_t|)$ auxiliary memory. 
At the end of the stream, the algorithm must output $S$ and $w$ such that  $\forall q \in \Q \; |F(q)  - \tilde F(q)| \le \eps m$ where 
$F(q) = \sum_{i=1}^{m} f(x_i,q)$ and $\tilde F(q) = \sum_{i \in S}w_i f(x_i,q)$.
The size of the streaming coreset is $\max_t |S_t|$.
\end{definition}

\begin{definition} [\bf Streaming Coreset Complexity] 
The Streaming Coreset Complexity for $f:\X,\Q \rightarrow \R$ is the minimal streaming coreset size among all possible 
streaming coreset algorithms for $f$.
\end{definition}

@@

\begin{theorem} \label{thm:streaming11}
Any function $f$ with Class Discrepancy $K_m(f) = O(c/m)$ has Streaming Coreset Complexity of $O\left(c\log^2(\eps n/c)/\eps\right)$.
The constant $c$ does not depend on $m$.
\end{theorem}

\begin{theorem} \label{thm:streaming21}
Any function $f$ with Class Discrepancy $K_m = O(c/\sqrt{m})$ has Streaming Coreset Complexity of $O\left(c^2\log^3(\eps^2 n/c) /\eps^2\right)$.
The constant $c$ does not depend on $m$.
\end{theorem}

\begin{theorem} \label{thm:streaming12}
Any function $f$ with Class Discrepancy $K_m(f) = O(c/m)$ has Streaming Coreset Complexity of $O\left(c\log^2\log(|Q_\eps|/\delta)/\eps\right)$.
The constant $c$ does not depend on $m$, $Q_\eps$ is an epsilon net for $f$ on $\Q$. 
The streaming coreset algorithm is randomized and fails with probability at most $\delta$.
\end{theorem}

\begin{theorem} \label{thm:streaming22}
Any function $f$ with Class Discrepancy $K_m(f) = O(c/\sqrt{m})$ has Streaming Coreset Complexity of $O\left(c^2\log^3\log(|Q_\eps|/\delta) /\eps^2\right)$.
The constant $c$ does not depend on $m$, $Q_\eps$ is an epsilon net for $f$ on $\Q$. 
The streaming coreset algorithm is randomized and fails with probability at most $\delta$.
\end{theorem}

@@

\begin{lemma} \label{lem:komlos anl}
Let $f$ be a function of the inner product $f(x,q) = f(\langle x,q\rangle)$ and let $f = \sum_k \alpha_k \langle x,q\rangle^k$ be its Taylor expansion. 
The discrepancy of $f$ indexed by $\|q\| \leq 1$ is bounded by
\[
K_m = \min_\sigma \sum_i \sigma_i f(x_i,q) =O\left( \sqrt{d} \sum_k  |\alpha_k|\sqrt{k\log^3(k)}\right)
\]
For general $\|q\| \leq R$ we get
\[
K_m = \min_\sigma \sum_i \sigma_i f(x_i,q) =O\left( \sqrt{d} \sum_k  |\alpha_k| R^k \sqrt{ k\log^3(k)}\right)
\]
\end{lemma}
\begin{proof}
The proof follows from combining the above.
$$
\sum_i \sigma_i f(x_i,q) = \sum_k \alpha_k \sum_i \sigma_i \ip{ x_i,q}^k =  \sum_k \alpha_k  \ip{  \sum_i \sigma_i x_i^{\otimes k},q^{\otimes k}} \le $$
$$\sum_k |\alpha_k| \cdot \left\| \sum_i \sigma_i x_i^{\otimes k}\right\|_{T_k} \cdot \|q\|^k
$$
By Lemma \ref{uc} we can find signs $\sigma$ such that 
$\left\| \sum_i \sigma_i x_i^{\otimes k}\right\|_{T_k} \le c\sqrt{d k \log^3(k)}$. Substituting into the above, the lemma follows.
\end{proof}

\begin{theorem}\label{analitic1}
Let $f:\R\rightarrow\R$ be analytic. There exist a radius $R$ such that the function family of functions $f = f(\ip{q,x})$, indexed by $\|q\| \leq R$, has discrepancy $O(\sqrt{d}/m)$. 
\end{theorem}
\begin{proof}
Recall that for analytic functions $f$ we have $\left| \frac{d^k f}{dz^k}(z) \right|  \leq C^{k+1} k! $
for some constant $C$. Considering the taylor expansion of $f$ near zero, for $R < 1/C$ the sum
$ \sum_k  |\alpha_k| R^k \sqrt{ k\log^3(k)} \leq C \sum_k  (CR)^k \sqrt{ k\log^3(k)}$
corresponding to Lemma~\ref{lem:komlos anl} converges to a constant. The result follows.
\end{proof}

\begin{corollary}
The discrepancy of the Logistic function $f(\ip{q,x}) = \log(1+\exp(\ip{q,x}))$ in dimension $d$, for $\|q\| \leq 1$ is $O(\sqrt{d}/m)$.
\end{corollary}

\begin{corollary}
The discrepancy of the covariance function $f(\ip{q,x}) = \ip{q,x}^2$ in dimension $d$, for $\|q\| \leq 1$ is $O(\sqrt{d}/m)$. This gives coresets for matrix column subset selection such that $\|XX^T - \tilde X \tilde X^T\| \le \eps n$ where $\tilde X$ contains $O(\sqrt{d}/\eps)$ rescaled columns of the matrix $X$.
\end{corollary}

\begin{theorem} \label{thm:analytic2}
Let $f:\R\rightarrow\R$ be analytic. There exist a radius $R$ such that the function family of functions $f_q(x) = f(\|x-q\|^2)$, indexed by $\|q\| \leq R$, has discrepancy $O(\sqrt{d}/m)$. 
\end{theorem}
\begin{proof}
By transforming $x$ to $\tilde{x} = (1, \sqrt{2}x, \|x\|^2)$ and $q$ to $\tilde{q} = (\|q\|^2, -\sqrt{2}q, 1)$ we get $\ip{\tilde{x},\tilde{q}} = \|q-x\|^2$. Moreover, $\|q\| \le R$ gives $\|\tilde q\| \le R^2+1$. The result follows from applying Theorem~\ref{analitic1} to $f(\ip{ \tilde q, \tilde x}) = f(\|q-x\|^2)$.
\end{proof}

\begin{corollary}
The discrepancy of the Gaussian kernel $K(q,x) = \exp(-\gamma \|x-q\|^2)$ in dimension $d$ is $O((1+\gamma)\exp(\gamma)\sqrt{d}/m)$.
\end{corollary} 
This improves upon the recent result of \cite{DBLP:journals/corr/abs-1802-01751} by proving the existence of $\eps$ approximation corsets of size $\sqrt{d}/\eps$ for Gaussian kernel density, in the case where $\gamma$ is constant. 
This also resolves the open problem raised by \cite{DBLP:journals/corr/abs-1802-01751} and matches their lower bound.

@@

\section{A Simple Algorithm for Kernel Density Estimation}
From section \ref{sec:analytic} we know that the discrepancy of the Gaussian kernel is $K_m = O(\sqrt{d}/m)$. 
%The result however is non-constructive in the sense that we are not aware of an algorithm that can obtain the $\sigma$ signs efficiently. 
Here, we provide a computationally efficient bound that can be achieved with a straightforward algorithm of complexity $O(m^2)$. Together with the results of Section~\ref{sec:sketch} this provides an efficient sketching algorithm for Kernel Density Estimation. 
In fact, we show that for any positive kernel $K_m = O(1/\sqrt{m})$. This bound is superior to that of the previous section for high dimensions $d > \sqrt{m}$. More importantly though, there is a very simple, intuitive, and deterministic algorithm for computing the signs $\sigma$. 
Given a collection of data points $X = x_1,\ldots, x_n$ in $\R^d$ the density function $f: \R^d \rightarrow \R$ of a point $q$ is defined as 
$$ F(q) = \sum_{i=1}^{n} K(x_i,q) $$
Here, $K$ is a \emph{positive definite kernel} function, typically based on the distance between $x,y$. The most frequent examples include
$$ K(x,q) = \exp(- \|x-y\|_2^2/\lambda^2)\;\;\; K(x,q) = \exp(- \|x-q\|/\lambda) \; \mbox{and}\;\;\; K(x,y) = (1+\|x-q\|/_2^2/\lambda^2)^{-1}$$
where $\lambda$ is a scaling parameter. For simplicity, we assume that $K(x,x) \leq 1$ for all data points. Notice that for any kernel based on distance we have $K(x,x)=1$ exactly for all $x \in \R^d$.

@@

The proofs of Theorems \ref{thm:streaming11}, \ref{thm:streaming12}, \ref{thm:streaming21}, and \ref{thm:streaming22} all use the basic concept of a compactor. A compactor consumes a stream of items and outputs another stream. 
The output stream contains at most half the items from the input stream with double the weight. 
It does so by keeping a buffer of a certain capacity $m$. When a new item is inserted into the compactor it is added to its buffer. 
If the buffer is full, a compaction operation takes place. 
The compaction takes the elements in the buffer $x_1,\ldots,x_m$ and finds a low discrepancy assignment $\sigma$ such that 
$\max_q |\sum_i \sigma_i f(x_i,q)| \leq m K_m$. 
Note that such a sequence is guarantied to exist by the definition of the discrepancy. For cases where an algorithm for finding this sequence $\sigma$ is not known, our result applies for the guarantee of the $\sigma$ sequence obtained by the algorithm. That is, if it is possible to obtain a bound of $K_m$ yet we can only find signs obtaining a bound of $\tilde{K}_m$, our results for the obtainable signs apply for $\tilde{K}_m$.
Given the sign vector $\sigma$, the compactor appends either $\{ x_i | \sigma_i = 1\}$ or  $\{ x_i | \sigma_i = -1\}$ to the output stream.

@@

Lemma~\ref{lem:det compactor} alone already allow us to prove Theorems~\ref{thm:streaming11} and~\ref{thm:streaming21}. 
The algorithms are a direct extension the well know MRL algorithm \cite{MRL} for quantile sketching. 
Note that for quantiles, $f(x,q) = 1$ if $q > x$ and $0$ else. 
A low discrepancy sequence is achieved simply by sorting the values and assigning $\sigma_i = 1$ for all evenly positioned values in the sorted order and $\sigma_i=-1$ to the odd positions. The above gives a discrepancy of $1/m$ for quantile approximation.
Below we generalize this algorithm to any low discrepancy class.

\paragraph{Theorem~\ref{thm:streaming11}}
For any function family $\F$ with a corresponding discrepancy $K_m = O(c/m)$ there exists an fully-mergeable streaming coreset deterministic algorithm of size $O\left(c\log^2(\eps n/c)/\eps\right)$ whose error is at most $\eps n$.
\begin{proof}
Consider feeding the output of the first compactor into a second one etc. Specifically, we start with a single compactor and open a second once it produced any output, then open a third compactor once the second produced output, etc.
Number the compactors $0,\ldots,H$. The weight of items given to compactors $h$ have weight $w_h = 2^h$. The length of the input stream seen by compactor is $n_h \le n/2^h$.

@@

\paragraph{Theorem~\ref{thm:streaming21}}
For any function family $\F$ with a corresponding discrepancy $K_m = O(c/\sqrt{m})$ there exists a fully-mergeable streaming coreset deterministic algorithm of size $O\left(c^2\log^3(\eps^2 n/c) /\eps^2\right)$ whose error is at most $\eps n$. 
\begin{proof}
The proof is identical to the one above except for the variable setting of
Setting $m \ge m_0 = O(c^2\log^2(\eps^2 n/c^2)/\eps^2)$ and replacing $K_m = c/\sqrt{m}$. 
We get that the error is at most $O(log(n/m) n K_m) \le \eps n$ for appropriate constants in the big-$O$ term of $m_0$, as required. 
Since we have $H = O(log(\eps^2 n/c^2))$ such compactors the overall space complexity is $O\left(c^2\log^3(\eps^2 n/c^2) /\eps^2\right)$. 
\end{proof}

@@

\paragraph{Theorem~\ref{thm:streaming22}}
For any function family $\F$ with a corresponding discrepancy $K_m = O(c/\sqrt{m})$ there exists a fully-mergeable streaming coreset randomized algorithm of size $O\left(c^2\log^3\log(n/\delta) /\eps^2\right)$ whose error for any fixed function $f \in \F$ is at most $\eps n$ with probability at least $1-\delta$. 
\begin{proof}
As in the deterministic setting we maintain a sequence of compactors of levels $h=0,\ldots,H$. Notice that the value of $H$ is increasing as the stream grows longer. Recall that a compactor of level $h$ observes elements of weight $2^h$ and outputs elements of weight $2^{h+1}$. As before we use a buffer of $m$ and get that $H \leq  \floor{log_2(n/m)}+1$.
The difference is that for a compactor of level $h$, once $h \leq H' = H - \log(\tilde{n}/m)$, where $\tilde{n} = O(\log(1/\delta)/\eps^2)$ with a constant in the $O()$ term that will be determined later, we change the mode of operation for this compactor. Notice that the requirement for $h$ ensures that the number of items observed by the $h$'th compactor is at least $n_h \geq \tilde{n}$. Rather than using a buffer of size $m$ the compactor uses a buffer of size 2 and for every two observed items it outputs one of them uniformly at random.

>>>>

\section{Introduction}
In both machine learning and in streaming and sketching problems our goal is (often) to approximate sums or expectations of well behaved functions.
Specifically, we need to approximate $\E_x f(x)$ or $\frac{1}{n}\sum_{i=1}^{n} f(x_i)$ for every $f\in \F$ where $\F$ is a family of functions and $x_i \in \X$ are either sampled training examples or an arbitrary set of stream items. 
Standard generalization results show that for a large enough value of $n$ the average approximates the mean if the complexity of $\F$ is bounded and the samples $x_i$ are drawn i.i.d.\ from an underlying distribution. We therefore focus on approximating the average, or rather, the sum $\sum_{i=1}^{n} f(x_i)$. For notational convenience we use a parameter $q \in \Q$ to index into $\F$ explicitly. 
In other words, there is a bijective mapping between $\Q \equiv \F$ such that $f(x) \in \F$ iff there exists $q\in\Q$ such that $f(x,q)=f(x)$.
We keep using the two different functions $f:\X\rightarrow\R$ and $f:\X,\Q\rightarrow\R$ interchangeably. 
One should think about $q$ as either the model parameters or a query for the sketch.

The goal is to produce a coreset. This is a small set $S\subset [n]$ and weights $w \in\R_+^{n}$ such that $\tilde F(q) = \sum_{i \in S} w_i f(x_i,q)$ approximates $F(q)$.
Approximation here means that $|\tilde F(q)  - F(q)| \le \eps n$ either for all $q \in \mathcal Q$ simultaneously. % or for every fixed $q$ with probability at least $1-\delta$. 
There are more complicated formulations such as weak coresets which we will not touch upon in this manuscript. 
Generating a concise representation $\tilde F$ for $F$ allows one to optimize over $\tilde F$ instead of $F$ which is more efficient. 
Moreover, if the resulting coresets are mergeable, this could be done on separate streams without the need for communication or assuming randomness in the partitioning.

@@

\subsection{Coreset Complexity}
In this section we point out a direct connection between coreset complexity and class discrepancy.
The connection is a simple application of the folklore argument know as the ``the halving trick".
For simplicity, in what follows we focus on functions $f$ whose range is $[0,1]$.

\begin{definition} [\bf Coreset Complexity] 
For a function $f:\X,\Q\rightarrow \R$ let $F(q) = \sum_{i=1}^{m} f(x_i,q)$ for a any set $\{x_1,\ldots,x_m\} \subset \X$.
For a set $S \subset [m]$ let $\tilde F(q) = \sum_{i \in S}w_i f(x_i,q)$ for some $w\in\R_+^m$ which is independent of $q$.
The coreset complexity of $f$ is the size of the smallest set $S$ such that $\forall q \in \Q \; |F(q)  - \tilde F(q)| \le \eps m$.
\end{definition}

\noindent The following facts are true for the common cases where $K_m = O(c/m)$ or $K_m = O(c/\sqrt{m})$.
\begin{fact}
Any function $f$ with class discrepancy $K_m(f) = O(c/m)$ has coreset complexity of $f$ $O(c/\eps)$.
The constant $c$ does not depend on $m$.
\end{fact}
\begin{fact}
Any function $f$ with class discrepancy $K_m(f) = O(c/\sqrt{m})$ has coreset complexity $O(c^2/\eps^2)$.
The constant $c$ does not depend on $m$.
\end{fact}
\begin{proof}
\noindent  Consider the signed sum of error function $E(q) = \sum_{i=1}^{m} \sigma_i f(x_i,q)$ where $\sigma_i \in \{-1,1\}$.
Now consider $\tilde F_{+}(q) = F(q) + E(q)   = \sum_{i | \sigma_i=1} 2 f(x_i,q)$ and similarly $\tilde F_{-}(q) = F(q) - E(q)  =  \sum_{i | \sigma_i=-1} 2 f(x_i,q)$. We have that both $\tilde F_{+}(q)$ and $\tilde F_{-}(q)$ are approximations for $F(q)$ obtained by coresets of item weights of $2$ and an error at most $|\tilde F_{\pm}(q)- F(q)| =  |E(q)|$. The above is true for any choice of signs $\sigma$, specifically, for those minimizing $\max_q | E(q)|$.
By Definition we can select signs such that $|E(q)| \le m K_m(f)$.

@@

\subsection{Streaming Coreset Complexity}\label{sec:sketch}
We claim that low class discrepancy implies concise streaming mergeable coresets as well. 
%
\begin{definition} [\bf Streaming Coreset Algorithm] 
A streaming coreset algorithm for $f:\X,\Q \rightarrow \R$ receives and arbitrary set $\{x_1,\ldots,x_m\} \subset \X$ one item after the other.
At time $t \le m$, the algorithm maintains a subset $S_t \subset \{x_1,\ldots,x_t\}$ and uses at most $O(|S_t|)$ auxiliary memory. 
At the end of the stream, the algorithm must output $S$ and $w$ such that  $\forall q \in \Q \; |F(q)  - \tilde F(q)| \le \eps m$ where 
$F(q) = \sum_{i=1}^{m} f(x_i,q)$ and $\tilde F(q) = \sum_{i \in S}w_i f(x_i,q)$.
The size of the streaming coreset is $\max_t |S_t|$.
\end{definition}

\begin{definition} [\bf Streaming Coreset Complexity] 
The streaming coreset complexity for $f:\X,\Q \rightarrow \R$ is the minimal streaming coreset size among all possible 
streaming coreset algorithms for $f$.
\end{definition}

@@

\begin{theorem} \label{thm:streaming11}
Any function $f$ with class discrepancy $K_m(f) = O(c/m)$ has streaming coreset complexity of $O\left(c\log^2(\eps n/c)/\eps\right)$.
The constant $c$ does not depend on $m$.
\end{theorem}

\begin{theorem} \label{thm:streaming21}
Any function $f$ with class discrepancy $K_m = O(c/\sqrt{m})$ has streaming coreset complexity of $O\left(c^2\log^3(\eps^2 n/c) /\eps^2\right)$.
The constant $c$ does not depend on $m$.
\end{theorem}

\begin{theorem} \label{thm:streaming12}
Any function $f$ with class discrepancy $K_m(f) = O(c/m)$ has streaming coreset complexity of $O\left(c\log^2\log(|Q_\eps|/\delta)/\eps\right)$.
The constant $c$ does not depend on $m$, $Q_\eps$ is an epsilon net for $f$ on $\Q$. 
The streaming coreset algorithm is randomized and fails with probability at most $\delta$.
\end{theorem}

\begin{theorem} \label{thm:streaming22}
Any function $f$ with class discrepancy $K_m(f) = O(c/\sqrt{m})$ has streaming coreset complexity of $O\left(c^2\log^3\log(|Q_\eps|/\delta) /\eps^2\right)$.
The constant $c$ does not depend on $m$, $Q_\eps$ is an epsilon net for $f$ on $\Q$. 
The streaming coreset algorithm is randomized and fails with probability at most $\delta$.
\end{theorem}

@@

\begin{lemma} \label{lem:komlos anl}
Let $f$ be a function of the inner product $f(x,q) = f(\langle x,q\rangle)$ and let $f = \sum_k \alpha_k \langle x,q\rangle^k$ be its Taylor expansion. 
The class discrepancy of $f$ indexed by $\|q\| \leq 1$ is bounded by
\[
K_m = \min_\sigma \sum_i \sigma_i f(x_i,q) =O\left( \sqrt{d} \sum_k  |\alpha_k|\sqrt{k\log^3(k)}\right)
\]
For general $\|q\| \leq R$ we get
\[
K_m = \min_\sigma \sum_i \sigma_i f(x_i,q) =O\left( \sqrt{d} \sum_k  |\alpha_k| R^k \sqrt{ k\log^3(k)}\right)
\]
\end{lemma}
\begin{proof}
The proof follows from combining the above.
$$
\sum_i \sigma_i f(x_i,q) = \sum_k \alpha_k \sum_i \sigma_i \ip{ x_i,q}^k =  \sum_k \alpha_k  \ip{  \sum_i \sigma_i x_i^{\otimes k},q^{\otimes k}} \le $$
$$\sum_k |\alpha_k| \cdot \left\| \sum_i \sigma_i x_i^{\otimes k}\right\|_{T_k} \cdot \|q\|^k
$$
By Lemma \ref{uc} we can find signs $\sigma$ such that 
$\left\| \sum_i \sigma_i x_i^{\otimes k}\right\|_{T_k} \le c\sqrt{d k \log^3(k)}$. Substituting into the above, the lemma follows.
\end{proof}

\begin{theorem}\label{analitic1}
Let $f:\R\rightarrow\R$ be analytic. There exist a radius $R$ such that the function family of functions $f = f(\ip{q,x})$, indexed by $\|q\| \leq R$, has class discrepancy $O(\sqrt{d}/m)$. 
\end{theorem}
\begin{proof}
Recall that for analytic functions $f$ we have $\left| \frac{d^k f}{dz^k}(z) \right|  \leq C^{k+1} k! $
for some constant $C$. Considering the taylor expansion of $f$ near zero, for $R < 1/C$ the sum
$ \sum_k  |\alpha_k| R^k \sqrt{ k\log^3(k)} \leq C \sum_k  (CR)^k \sqrt{ k\log^3(k)}$
corresponding to Lemma~\ref{lem:komlos anl} converges to a constant. The result follows.
\end{proof}

\begin{corollary}
The class discrepancy of the Logistic function $f(\ip{q,x}) = \log(1+\exp(\ip{q,x}))$ in dimension $d$, for $\|q\| \leq 1$ is $O(\sqrt{d}/m)$.
\end{corollary}

\begin{corollary}
The class discrepancy of the covariance function $f(\ip{q,x}) = \ip{q,x}^2$ in dimension $d$, for $\|q\| \leq 1$ is $O(\sqrt{d}/m)$. This gives coresets for matrix column subset selection such that $\|XX^T - \tilde X \tilde X^T\| \le \eps n$ where $\tilde X$ contains $O(\sqrt{d}/\eps)$ rescaled columns of the matrix $X$.
\end{corollary}

\begin{theorem} \label{thm:analytic2}
Let $f:\R\rightarrow\R$ be analytic. There exist a radius $R$ such that the function family of functions $f_q(x) = f(\|x-q\|^2)$, indexed by $\|q\| \leq R$, has class discrepancy $O(\sqrt{d}/m)$. 
\end{theorem}
\begin{proof}
By transforming $x$ to $\tilde{x} = (1, \sqrt{2}x, \|x\|^2)$ and $q$ to $\tilde{q} = (\|q\|^2, -\sqrt{2}q, 1)$ we get $\ip{\tilde{x},\tilde{q}} = \|q-x\|^2$. Moreover, $\|q\| \le R$ gives $\|\tilde q\| \le R^2+1$. The result follows from applying Theorem~\ref{analitic1} to $f(\ip{ \tilde q, \tilde x}) = f(\|q-x\|^2)$.
\end{proof}

\begin{corollary}
The class discrepancy of the Gaussian kernel $K(q,x) = \exp(-\gamma \|x-q\|^2)$ in dimension $d$ is $O((1+\gamma)\exp(\gamma)\sqrt{d}/m)$.
\end{corollary} 
This improves upon the recent result of \cite{DBLP:journals/corr/abs-1802-01751} by proving the existence of $\eps$ approximation corsets of size $\sqrt{d}/\eps$ for Gaussian kernel density, in the case where $\gamma$ is constant. 
This also resolves the open problem raised by \cite{DBLP:journals/corr/abs-1802-01751} and matches their lower bound.

@@

\section{A Simple Algorithm for Kernel Density Estimation}
From section \ref{sec:analytic} we know that the class discrepancy of the Gaussian kernel is $K_m = O(\sqrt{d}/m)$. 
%The result however is non-constructive in the sense that we are not aware of an algorithm that can obtain the $\sigma$ signs efficiently. 
Here, we provide a computationally efficient bound that can be achieved with a straightforward algorithm of complexity $O(m^2)$. Together with the results of Section~\ref{sec:sketch} this provides an efficient sketching algorithm for Kernel Density Estimation. 
In fact, we show that for any positive kernel $K_m = O(1/\sqrt{m})$. This bound is superior to that of the previous section for high dimensions $d > \sqrt{m}$. More importantly though, there is a very simple, intuitive, and deterministic algorithm for computing the signs $\sigma$. 
Given a collection of data points $X = x_1,\ldots, x_n$ in $\R^d$ the density function $f: \R^d \rightarrow \R$ of a point $q$ is defined as 
$$ F(q) = \sum_{i=1}^{n} K(x_i,q) $$
Here, $K$ is a \emph{positive definite kernel} function, typically based on the distance between $x,y$. The most frequent examples include
$$ K(x,q) = \exp(- \|x-y\|_2^2/\lambda^2)\;\;\; K(x,q) = \exp(- \|x-q\|/\lambda) \; \mbox{and}\;\;\; K(x,y) = (1+\|x-q\|/_2^2/\lambda^2)^{-1}$$
where $\lambda$ is a scaling parameter. For simplicity, we assume that $K(x,x) \leq 1$ for all data points. Notice that for any kernel based on distance we have $K(x,x)=1$ exactly for all $x \in \R^d$.

@@

The proofs of Theorems \ref{thm:streaming11}, \ref{thm:streaming12}, \ref{thm:streaming21}, and \ref{thm:streaming22} all use the basic concept of a compactor. A compactor consumes a stream of items and outputs another stream. 
The output stream contains at most half the items from the input stream with double the weight. 
It does so by keeping a buffer of a certain capacity $m$. When a new item is inserted into the compactor it is added to its buffer. 
If the buffer is full, a compaction operation takes place. 
The compaction takes the elements in the buffer $x_1,\ldots,x_m$ and finds a low discrepancy assignment $\sigma$ such that 
$\max_q |\sum_i \sigma_i f(x_i,q)| \leq m K_m$. 
Note that such a sequence is guarantied to exist by the definition of the class discrepancy. For cases where an algorithm for finding this sequence $\sigma$ is not known, our result applies for the guarantee of the $\sigma$ sequence obtained by the algorithm. That is, if it is possible to obtain a bound of $K_m$ yet we can only find signs obtaining a bound of $\tilde{K}_m$, our results for the obtainable signs apply for $\tilde{K}_m$.
Given the sign vector $\sigma$, the compactor appends either $\{ x_i | \sigma_i = 1\}$ or  $\{ x_i | \sigma_i = -1\}$ to the output stream.

@@

Lemma~\ref{lem:det compactor} alone already allow us to prove Theorems~\ref{thm:streaming11} and~\ref{thm:streaming21}. 
The algorithms are a direct extension the well know MRL algorithm \cite{MRL} for quantile sketching. 
Note that for quantiles, $f(x,q) = 1$ if $q > x$ and $0$ else. 
A low discrepancy sequence is achieved simply by sorting the values and assigning $\sigma_i = 1$ for all evenly positioned values in the sorted order and $\sigma_i=-1$ to the odd positions. The above gives class discrepancy of $1/m$ for quantile approximation.
Below we generalize this algorithm to any low discrepancy class.

\todo{Replace with the above?}
\paragraph{Theorem~\ref{thm:streaming11}}
For any function family $\F$ with a corresponding class discrepancy $K_m = O(c/m)$ there exists an fully-mergeable streaming coreset deterministic algorithm of size $O\left(c\log^2(\eps n/c)/\eps\right)$ whose error is at most $\eps n$.
\begin{proof}
Consider feeding the output of the first compactor into a second one etc. Specifically, we start with a single compactor and open a second once it produced any output, then open a third compactor once the second produced output, etc.
Number the compactors $0,\ldots,H$. The weight of items given to compactors $h$ have weight $w_h = 2^h$. The length of the input stream seen by compactor is $n_h \le n/2^h$.

@@

\paragraph{Theorem~\ref{thm:streaming21}}
For any function family $\F$ with a corresponding class discrepancy $K_m = O(c/\sqrt{m})$ there exists a fully-mergeable streaming coreset deterministic algorithm of size $O\left(c^2\log^3(\eps^2 n/c) /\eps^2\right)$ whose error is at most $\eps n$. 
\begin{proof}
The proof is identical to the one above except for the variable setting of
Setting $m \ge m_0 = O(c^2\log^2(\eps^2 n/c^2)/\eps^2)$ and replacing $K_m = c/\sqrt{m}$. 
We get that the error is at most $O(log(n/m) n K_m) \le \eps n$ for appropriate constants in the big-$O$ term of $m_0$, as required. 
Since we have $H = O(log(\eps^2 n/c^2))$ such compactors the overall space complexity is $O\left(c^2\log^3(\eps^2 n/c^2) /\eps^2\right)$. 
\end{proof}

@@

\paragraph{Theorem~\ref{thm:streaming22}}
For any function family $\F$ with a corresponding class discrepancy $K_m = O(c/\sqrt{m})$ there exists a fully-mergeable streaming coreset randomized algorithm of size $O\left(c^2\log^3\log(n/\delta) /\eps^2\right)$ whose error for any fixed function $f \in \F$ is at most $\eps n$ with probability at least $1-\delta$. 
\begin{proof}
As in the deterministic setting we maintain a sequence of compactors of levels $h=0,\ldots,H$. Notice that the value of $H$ is increasing as the stream grows longer. Recall that a compactor of level $h$ observes elements of weight $2^h$ and outputs elements of weight $2^{h+1}$. As before we use a buffer of $m$ and get that $H \leq  \floor{log_2(n/m)}+1$.
The difference is that for a compactor of level $h$, once $h \leq H' = H - \log(\tilde{n}/m)$, where $\tilde{n} = O(\log(1/\delta)/\eps^2)$ with a constant in the $O()$ term that will be determined later, we change the mode of operation for this compactor. Notice that the requirement for $h$ ensures that the number of items observed by the $h$'th compactor is at least $n_h \geq \tilde{n}$. Rather than using a buffer of size $m$ the compactor uses a buffer of size 2 and for every two observed items it outputs one of them uniformly at random.
